---
title: "Problem Set 1"
author: "Prof. Bisbee"
institute: "Vanderbilt University"
date: "Due Date: 2022/09/11 @ 6PM CST"
output:
  html_document:
    self_contained: yes
  pdf_document: default
---

## Getting Set Up

If you haven't already, create a folder for this course, and then a subfolder within for the third week `Topic3_HelloWorld`, and two additional subfolders within `code` and `data`.

Open `RStudio` and create a new RMarkDown file (`.Rmd`) by going to `File -> New File -> R Markdown...`.
Accept defaults and save this file as `[LAST NAME]_ps1.Rmd` to your `code` folder.

Copy and paste the contents of this `Topic3_HelloWorld_PS1.Rmd` file into your `[LAST NAME]_ps1.Rmd` file. Then change the `author: [Your Name]` to your name.

If you haven't already, download the `sc_debt.Rds` file from the course [github page](https://github.com/jbisbee1/DSCI1000/blob/main/Lectures/Topic3_HelloWorld/data/sc_debt.Rds) and save it to your `data` folder.

All of the following questions should be answered using 
````{verbatim}
```{r}
# EXAMPLE CODE
mean(c(1,2,3))
```
````

Some questions will also ask for answers written out in English, instead of just `R` code. These will be indicated with the code

````{verbatim}
> - Write answer here
````


This problem set is worth 20 total points, plus one extra credit point. The point values for each question are indicated in brackets below. You are free to rely on whatever resources you need to complete this problem set, including lecture notes, lecture presentations, Google, your classmates...you name it. However, the final submission must be complete by you. There are no group assignments. To submit, compiled the completed problem set and upload the `.html` file to Brightspace by 6PM CST on 2022/09/11.

**Good luck!**


## Question 1 [1 point]
*Require `tidyverse` and load the `sc_debt.Rds` data by assigning it to an object named `df`.*
```{r}
# INSERT CODE HERE
```


## Question 2 [2 points]
*Which school has the lowest admission rate?*
```{r}
# INSERT CODE HERE
```

> - Write answer here (1-2 sentences)

## Question 3 [2 points]
*Which are the top 10 schools by average SAT score?*
```{r}
# INSERT CODE HERE
```

## Question 4 [2 points]
*Which state is home to the school with the largest median earnings of recent graduates, and how much did they make?*
```{r}
# INSERT CODE HERE
```

> - Write answer here (1-2 sentences)

## Question 5 [2 points]
*What is the average SAT score of the school with the highest median earnings identified in question 3?*
```{r}
# INSERT CODE HERE
```

> - Write answer here (1-2 sentences)

## Question 6 [2 points]
*Calculate the average SAT score and median earnings of recent graduates by state.*
```{r}
# INSERT CODE HERE
```

## Question 7 [2 points]
*Plot the average SAT score (x-axis) against the median earnings of recent graduates (y-axis) by school.*
```{r}
# INSERT CODE HERE
```

## Question 8 [4 points]
*Plot the average SAT score (x-axis) against the median earnings of recent graduates (y-axis) by **STATE**. (Extra credit: plot the line of best fit.)*
```{r}
# INSERT CODE HERE
```

## Question 9 [4 points]
*What relationship do you observe? Why do you think this relationship exists?*

> - Write answer here (3-4 sentences)

